---
title: "Reproducible Research: Peer Assessment 1"
author: "Sergi Marti"
output: 
  html_document:
    keep_md: true
---

## Required Libraries

```{r, echo=TRUE, message=FALSE}
library(data.table)
library(ggplot2)

```

## Loading and preprocessing the data

```{r, echo=TRUE}

unzip(zipfile = "activity.zip")
data_activity <- data.table::fread(file = "activity.csv", sep = ",", dec = ".")


str(data_activity)
```


## What is mean total number of steps taken per day?

```{r, echo=TRUE}

#Total number of steps per day
data_hist <- data_activity[!(steps == 0) & !(is.na(steps)) 
                           ,.("steps" = sum(steps))
                           , by = date]


#Histogram
hist(data_hist$steps, main = "Histogram - Numbers of steps taken per day")

```


```{r, echo=TRUE}

#Mean and Meian
data_hist[,.("Mean"    = mean(steps)
             ,"Median" = median(steps))]

```



## What is the average daily activity pattern?

```{r, echo=TRUE}

data_patt <- data_activity[,.("steps" = mean(steps, na.rm = T)), interval][order(interval)]
plot(data_patt$interval
     , data_patt$steps
     , type = "l"
     , main = "Average daily activity pattern")

```

```{r, echo=TRUE}

#Interval that contain the maximum number of steps
data_patt[steps == max(steps), interval]

```


## Imputing missing values

```{r, echo=TRUE}

#Total number missing values
nrow(data_activity[is.na(steps)])

```


```{r, echo=TRUE}

data_activity_no_NA <- copy(data_activity)

#Filling NA
data_activity_no_NA[,steps := as.numeric(steps)]
data_activity_no_NA[is.na(steps), steps := mean(data_activity$steps, na.rm = TRUE)]

#Total number of steps per day
data_hist_no_NA <- data_activity_no_NA[ ,.("steps" = sum(steps))
                                       , by = date]

#Histogram
hist(data_hist_no_NA$steps, main = "Histogram - Numbers of steps taken per day (w/o NA)")

```



```{r, echo=TRUE}

#Mean and Meian
data_hist_no_NA[,.("Mean"    = mean(steps)
                   ,"Median" = median(steps))]


```

Imputing missing data, especially using the mean, is a common method for handling gaps in a dataset. However, it does have several implications and potential impacts on the estimates and future analyses. Here's a example of the effects:

  - Bias
  - Variance Reduction
  - Data appear more normally distributed than it actually is (overestimation of Data Quality)
  - Influence on Statistical Inferences
  - Inflated estimates
  - ...
  

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}

#Coverting date
data_activity_no_NA[, wd := weekdays(as.Date(date, "%Y-%m-%d"))]
data_activity_no_NA[, wd := ifelse(wd %like% "S", "Weekend", "Weekday")]
data_activity_no_NA[, wd := as.factor(wd)]

data_activity_no_NA_wd <- data_activity_no_NA[,.("steps" = mean(steps))
                                              , .(interval, wd)]

#Plot
p <- ggplot(data_activity_no_NA_wd, aes(x=interval, y=steps, group=wd, col=wd, fill=wd)) +
      geom_line() +
      facet_grid(rows = vars(wd))

p
```


